% Use of LuaTeX seems to be recommended.
% This stuff is all about "standard" LaTex capabilities used in the document.
\documentclass[
	techspec,	% for a Technical Specification
	final,		% needed for PDF bookmarks
	notcopyright,	% no copyright statements (yet)
	letterpaper	% for now
	]{isov2}
\let\ifpdf\relaxed	% let hyperref implement \ifpdf
\usepackage[
	pdftitle={Extensions for parallel programming},
	pdfauthor={Clark Nelson},
	pdfsubject={WORKING DRAFT Technical Specification},
	pdfkeywords={},
	pdfstartview=FitH,
	colorlinks=true, % not boxed
	linkcolor=blue,	% usual default for browsers
	urlcolor=blue
	]{hyperref}	% for PDF hyperlinks and bookmarks
\usepackage{chngcntr}	% for paragraph numbering
\usepackage{color}	% for editorial marking
\usepackage[T1]{fontenc} % to use 256-character fonts
\usepackage{fontspec}	% to disable ligatures
\usepackage{ifthen} 	% for conditional code/text
\usepackage{listings}	% for code listings
\usepackage{makeidx}	% for indexing
\usepackage[normalem]{ulem}	% for underlining and strikeout
\usepackage{underscore} % to simplify use of underscores
\usepackage{verbatim}	% for code font
\usepackage{xspace}	% to simplify macros that expand to text
\makeindex
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{2} % practical TOC size
\setcounter{tocdepth}{6} % uncomment to check outline structure

% These macros are from the WG21 standard source.
\input{macros}
\renewcommand{\indextext}[1]{\index{#1}}
\renewcommand{\indexgram}[1]{\index{#1}}

% This is for macros invented for/by CPLEX.
\newboolean{hidecpp}
\setboolean{hidecpp}{false}
\input{cplexmac}

% These parameters, and the documentclass options,
% are about document identification and overall formatting.
\newcommand{\cplexts}{$CPLEXTS$}
\renewcommand{\extrahead}{2015-08-31}
\standard{\cplexts}
\yearofedition{2017}
\languageofedition{(E)}

%\includeonly{reduction}

\begin{document}

\begin{cover}
\begin{tabular}{l l}
\textbf{Revision:} & \extrahead \\
\textbf{Reply to:} & Clark Nelson
\end{tabular}

{\Huge
\bfseries
Programming languages --- C --- \\
Extensions for parallel programming
}

This is the latest working draft of the CPLEX study group.
\clearpage
\end{cover}

% The table of contents appears here.

\begin{foreword}
\input{tspasfwdbp}

ISO/TS
\cplexts{}
was prepared by Technical Committee ISO/IEC JTC1/SC22/WG14.%
\footnote{FYI:
This is the only paragraph in the Foreword that has anything in it
that's not just boilerplate.
}

\fwdnopatents
\end{foreword}

\begin{introduction}
\intropatents
\end{introduction}

\title{Programming languages}{C}{Extensions for parallel programming}

\scopeclause
\begin{inscope}{technical specification}
\item
Extensions to the C language to simplify writing a parallel program.
\end{inscope}
\begin{outofscope}{technical specification}
\item
Support for writing a concurrent program.
\end{outofscope}

\normrefsclause
\normrefbp{technical specification}
\begin{nreferences}
\isref{ISO/IEC 9899:2011(E)}{Programming languages --- C}
\isref{ISO/IEC 14882:2014(E)}{Programming languages --- C++}
\end{nreferences}

\include{definitions}

\clause{Document conventions}

\pnum
\begin{cpp}
Text that is specific to C++
is enclosed in square brackets
and presented in oblique sans-serif type.
\end{cpp}

\pnum
Definitions of terms and grammar non-terminals defined in the C
\begin{cpp}
or C++
\end{cpp}
standard are not duplicated in this document.
Terms and grammar non-terminals defined in this document
are referenced in the index.
The ``cplex_'' prefix of library identifiers
is omitted from the index entry.

\pnum
According to the ISO editing directives,
the use of footnotes
``shall be kept to a minimum.''
Almost all of the footnotes in this document
are not intended to survive to final publication.
Most footnotes are classified by an abbreviation:
\begin{description}
\item[FYI]
A point of information.
\item[TODO]
Calls attention to an area that needs more work.
\item[DFEP]
Departure from existing practice.
\end{description}

\begin{comment}
\pnum
Annex A contains information
concerning the editing of the LaTeX source of this document.
It will not survive to final publication.
\end{comment}

\include{tasks}

\include{countable}

\include{parallelloop}

\include{spawning}

\include{reduction}

\include{loopparameter}

%\%include{samples}

\bibannex
\begin{references}
\reference{}{Intel\copyright{} Cilk\texttrademark{} Plus
Language Extension Specification,}
{Intel Corporation}:
\isourl{https://www.cilkplus.org/sites/default/files/open_specifications/Intel_Cilk_plus_lang_spec_1.2.htm}
\reference{}{OpenMP Application Program Interface,}
{OpenMP Architecture Review Board}:
\isourl{http://www.openmp.org/mp-documents/OpenMP4.0.0.pdf}
\end{references}

% we definitely want a multi-column index, but we don't get one by default
\twocolumn % even adding this doesn't get the job done
\printindex
\end{document}
