ISO (the International Organization for Standardization) is a worldwide federation of national standards
bodies (ISO member bodies). The work of preparing International Standards is normally carried out
through ISO technical committees. Each member body interested in a subject for which a technical
committee has been established has the right to be represented on that committee. International organizations,
governmental and non-governmental, in liaison with ISO, also take part in the work. ISO
collaborates closely with the International Electrotechnical Commission (IEC) on all matters of electrotechnical
standardization.

International Standards are drafted in accordance with the rules given in the ISO/IEC Directives, Part 2.

The main task of technical committees is to prepare International Standards. Draft International Standards
adopted by the technical committees are circulated to the member bodies for voting. Publication
as an International Standard requires approval by at least 75\% of the member bodies casting a vote.

In other circumstances, particularly when there is an urgent market requirement for such documents, a
technical committee may decide to publish other types of normative document:

\begin{itemize}
\item
an ISO Publicly Available Specification (ISO/PAS) represents an agreement between technical experts
in an ISO working group and is accepted for publication if it is approved by more than 50\%
of the members of the parent committee casting a vote;
\item
an ISO Technical Specification (ISO/TS) represents an agreement between the members of a technical
committee and is accepted for publication if it is approved by 2/3 of the members of the
committee casting a vote.
\end{itemize}

An ISO/PAS or ISO/TS is reviewed every three years with a view to deciding whether it can be transformed
into an International Standard.
